---
title: 'Lifestyle and Socio-demographic Factors Affecting Weight Variability Among Canadian Adult'
author: "Sena Okuboyejo"
date: "10/12/2021"
html_document:
      pdf_document: default

---



```{r Package Installation}
#install.packages("naniar")
#install.packages("visdat")
#install.packages("finalfit")
#install.packages("vtable")
#install.packages("readr")


```

```{r Load relevant libraries}
library(readr)
library(corrr)
library(nortest)
library(GGally)
library(Hmisc)
library(readxl)
library(psych)          
library(tidyverse)
library(janitor)
library(rstatix)
library(NHANES)
library(table1)
library(kableExtra)
library(car)
library(stargazer)
library(ggpubr)
library(pROC)
library(ggplot2)
library(gridExtra)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
library(dplyr)
library(gtsummary)
library(naniar)
library(visdat)
library(nnet)
library(pscl)
library(MASS)
library(caret)
library(leaps)
library(finalfit)
library(vtable)
library(readr)


knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, cache=TRUE,fig.width=8, fig.height=5)
options(full_width = FALSE)
options(knitr.kable.NA = '')

cchsData <- read_csv("~/DGIN5400/FinalProject/cchsData.csv")

```

```{r Summary Statistics of entire dataset}
dataset1 <- data.frame(cchsData) #View dataset


#str(dataset1) #View dataset

#dataset1%>% tbl_summary()

```

#Data Preparation and Cleaning

•	BMI–HWTDGBMI Outcome variable
•	Race – SDCDGCGT
•	Sleep at Night – SLPG005
•	Sedentary Behavior – SBE_005
•	Age - DHHGAGE
•	Food Choice - FDC_005
•	Gender (Sex) – DHH_SEX
•	
```{r Selecting Variables}
# choosing specific variables

lifestyleVar = c("hwtdgbmi","dhhgage","dhh_sex","sdcdgcgt", "slpg005","sbe_005",   "fdc_005" )

# create a new dataset with only the variables for the study

lifestyle_dat = dataset1[,lifestyleVar]
```
 

```{r Describe the data}
summary(lifestyle_dat)


```

#Data Cleaning

```{r Identify values of each variable}
str(lifestyle_dat)

```
```{r}
lifestyle_dat2 <- lifestyle_dat %<>% 
  
  # remove Under 18 (groups 1 and 2)
filter(dhhgage>2) %>%
  
 #Collapse age
  mutate(decade=case_when(
   dhhgage %in% 3:8 ~ 1,
    dhhgage %in% 9:12 ~ 2,
    dhhgage %in% 13:16 ~ 3
  ),
  
# make females reference i.e.  0
    dhh_sex=replace(dhh_sex,dhh_sex==2,0),
  
# Make race binary - White/Non-white
      race=case_when(
          sdcdgcgt %in% c(2) ~0,
          sdcdgcgt %in% c(1) ~1,
          sdcdgcgt > 3         ~NA_real_),

# Make food choices - body weight concerns a binary variable No/Yes
  foodChoice=case_when(
         fdc_005 %in% c(1)   ~1,
          fdc_005 %in% c(2) ~0,
          fdc_005 > 2 ~NA_real_),

# Make sleep hours at night a binary variable Insufficient sleep/ Sufficient sleep
    sleepHrs=case_when(
           slpg005 %in% c(1:6)~0,
           slpg005 %in% c(7:11)  ~1,
            slpg005 %in% c(96)  ~NA_real_),
         
# Make sedentary behaviouron weekdays levels
    sedBehave=case_when(
            sbe_005 %in% c(1,2)   ~0,
            sbe_005 %in% c(3,4) ~1,
            sbe_005 %in% c(5) ~2,
            sbe_005 %in% c(6) ~3,
            sbe_005 > 6 ~NA_real_)

  )
head(lifestyle_dat2)
  
    
          
         
          
```



```{r View dataset}
summary(lifestyle_dat2)
```


```{r Change to factors}
 # change to factors

lifestyle_dat3 = lifestyle_dat2 |> tibble() |>
  mutate(
    BMIGroups = cut(hwtdgbmi,breaks=c(0,18.5,25,30,Inf),labels=c('Underweight','Normal','Overweight','Obese')),
    Decade=factor(decade,levels=1:3,labels=c('Youth Group','Middle-aged Group','Elderly Group')),
    Sex = factor(dhh_sex,levels=0:1,labels=c('Female','Male')),
    Race = factor(race,levels=0:1,labels=c('Non-white','White')),
    FoodChoice=factor(foodChoice,levels=c(0:1), labels=c('No Weight Concerns', 'Yes -Weight Concerns')),
  SedentaryBehaviour =factor(sedBehave,levels=c(0,1,2,3), labels = c('0-4','4-7', '8+', 'Not at work')),
  SleepHrs=factor(sleepHrs,levels=c(0:1), labels = c('Insufficient sleep','Sufficient sleep')),

  
)

#adding labels 
label(lifestyle_dat3$Sex) = "Sex"
label(lifestyle_dat3$Decade) = "Age"
label(lifestyle_dat3$BMIGroups) = "Self-reported BMI"
label(lifestyle_dat3$Race) = "Race"
label(lifestyle_dat3$SleepHrs) = "Sleep hours at night"
label(lifestyle_dat3$SedentaryBehaviour) = "Time sitting /lying watching screen - school / workday"
label(lifestyle_dat3$FoodChoice) = "Choose food - Concerned about body weight"

#Creating reference levels for the categorical variables in the model
lifestyle_dat3$BMIGroups = relevel(lifestyle_dat3$BMIGroups,'Normal')
lifestyle_dat3$Sex = relevel(lifestyle_dat3$Sex,'Female')
lifestyle_dat3$Decade = relevel(lifestyle_dat3$Decade,ref='Youth Group')
lifestyle_dat3$Race= relevel(lifestyle_dat3$Race,ref='White')
lifestyle_dat3$SleepHrs = relevel(lifestyle_dat3$SleepHrs,ref='Sufficient sleep')
lifestyle_dat3$FoodChoice = relevel(lifestyle_dat3$FoodChoice,ref='No Weight Concerns')
lifestyle_dat3$SedentaryBehaviour = relevel(lifestyle_dat3$SedentaryBehaviour,ref='0-4')


str(lifestyle_dat3)
```

#Summary Statistics

```{r Summary Statistics}

summary(lifestyle_dat3)

```

#Check for missingness

```{r Check for Missingness}

# choosing specific variables

lifestyleVar1 = c("FoodChoice","SleepHrs","SedentaryBehaviour","Race", "Decade","BMIGroups", "Sex" )
lifestyleVar2 = c("FoodChoice","SleepHrs","SedentaryBehaviour","Race", "Decade","BMIGroups", "Sex","hwtdgbmi" )



# create a new dataset with only the variables you want
lifestyle_dat4 = lifestyle_dat3[,lifestyleVar1]
lifestyle_dat5 = lifestyle_dat3[,lifestyleVar2]

gg_miss_var(lifestyle_dat4, show_pct = TRUE)
gg_miss_var(lifestyle_dat5, show_pct = TRUE)
```


```{r Check complete cases with BMI as a factor variable}

lifestyle_dat4 <- lifestyle_dat4 %>% 
  drop_na()

summary(lifestyle_dat4)

```

```{r Check complete cases with BMI as continuous variable}
lifestyle_dat5 <- lifestyle_dat5 %>% 
  drop_na()

str(lifestyle_dat5)
```



```{r Summary Statistics of final variables}

lifestyle_dat4 %>%
  tbl_summary()

```


```{r Summary Statistics according to BMIGroups}

lifestyle_dat4  %>% 
  tbl_summary(by = BMIGroups) %>% 
  add_p()


```

#Export table

```{r}
lifestyle_summary = st(lifestyle_dat4, group = 'BMIGroups', group.test = TRUE, out = 'csv')
lifestyle_summary 

#tab_model(lifestyle_summary)
```

```{r}
write.table(lifestyle_summary, file = "lifestylesummary.txt", sep = "\t",
            row.names = TRUE, col.names = NA)
```

```{r}
st(lifestyle_dat4, group = 'BMIGroups', group.test = TRUE, out ="kable")

```



#Descriptive Statistics of Variables

```{r}
#p = lifestyle_dat4 |> 
  #select(FoodChoice,SleepHrs,SedentaryBehaviour,Race, Decade,BMIGroups, Sex) |>
  #pivot_longer(cols=-BMIGroups,names_to='var',values_to='Days') |>
  #group_by(BMIGroups,var,Days) |>
  #summarize(count=n())|>
  #ggplot(aes(y=count,x=Days,fill=BMIGroups))+
  #geom_bar(stat='identity',position='dodge')+
  #facet_wrap(~var)
#p
```
```{r Setting the reference}
lifestyle_dat4$BMIGroups1 <-relevel(lifestyle_dat4$BMIGroups, ref ="Normal")

#check that "Normal" is now the reference for BMIGroups

levels(lifestyle_dat4$BMIGroups1)
```

#Correlation Matrix

```{r Correlation matrix usinf continuous variables}
correlation.matrix <- cor(lifestyle_dat[,c("hwtdgbmi","dhhgage","dhh_sex","sdcdgcgt", "slpg005","sbe_005","fdc_005")])

correlation.matrix

#formatting the correlation matrix

stargazer(correlation.matrix, header = FALSE, type = "Text", title = "Correlation Matrix")
```


#Univariate Analysis



```{r Linear Regression}

modReg1 = lm(hwtdgbmi~Decade,data=lifestyle_dat5)
Anova(modReg1)
TukeyHSD(aov(modReg1))

modReg2 = lm(hwtdgbmi~Sex,data=lifestyle_dat5)
Anova(modReg2)
TukeyHSD(aov(modReg2))

modReg3 = lm(hwtdgbmi~SedentaryBehaviour,data=lifestyle_dat5)
Anova(modReg3)
TukeyHSD(aov(modReg3))

modReg4 = lm(hwtdgbmi~FoodChoice,data=lifestyle_dat5)
Anova(modReg4)
TukeyHSD(aov(modReg4))

modReg5 = lm(hwtdgbmi~SleepHrs,data=lifestyle_dat5)
Anova(modReg5)
TukeyHSD(aov(modReg5))

modReg6 = lm(hwtdgbmi~Race,data=lifestyle_dat5)
Anova(modReg6)
TukeyHSD(aov(modReg6))


```

#Formatting Linear Regression results using Stargazer

```{r}
stargazer(modReg1, modReg2, modReg3, modReg4, modReg5, modReg6, type = "text", title = "Linear Models")
```
```{r}
ologreg_lifestyle = glm( BMIGroups1~Sex + FoodChoice + SleepHrs + SedentaryBehaviour+  Race + Decade,family=binomial, data = lifestyle_dat4)

Anova(ologreg_lifestyle)
summary(ologreg_lifestyle)
```


#Multinomial Logistic Regression of all variables without interaction

```{r}
logreg_lifestyle = multinom( BMIGroups1~Sex + FoodChoice + SleepHrs + SedentaryBehaviour+  Race + Decade, data = lifestyle_dat4)

Anova(logreg_lifestyle)
summary(logreg_lifestyle)
```




#Model Fitting

#Calculating the fit using the McFadden's Pseudo R2
# Here, we will use McFadden’s Pseudo R2. The function, pR2(), is included in the pscl package by Simon Jackman at Stanford University, which you can download using the install.packages() command.

```{r}
# Get the McFadden pR2 from our logistic model
pR2(logreg_lifestyle)
```


#Computing stepwise regression

stepAIC() [MASS package], which choose the best model by AIC. It has an option named direction, which can take the following values: i) “both” (for stepwise regression, both forward and backward selection); “backward” (for backward selection) and “forward” (for forward selection). It returns the best final model.

```{r}


# Stepwise regression model
step.model <- stepAIC(logreg_lifestyle, direction = "both", trace = FALSE)

summary(step.model)

```


#Final model

```{r}
explanatory = c("FoodChoice", "Sex", 
  "Decade", "SleepHrs", "SedentaryBehaviour", "Race")
dependent = "BMIGroups"

lifestyle_dat4 %>% 
  finalfit(dependent, explanatory, 
  dependent_label_prefix = "") -> table2
table2


```


#Logistic Regression of all variables while checking for interaction between sex and other variables

```{r}
logreg_lifestyle_sex = multinom( BMIGroups1~Sex*FoodChoice + Sex*SleepHrs + Sex*SedentaryBehaviour+  Sex*Race + Sex*Decade, data = lifestyle_dat4)

Anova(logreg_lifestyle_sex)
summary(logreg_lifestyle_sex)
```

#Logistic Regression of all variables while removing non-significant interactions - sex:race, sex:sleephrs
```{r}
logreg_lifestyle_sex1 = multinom( BMIGroups1~Sex*FoodChoice + SleepHrs + Sex*SedentaryBehaviour+  Race + Sex*Decade, data = lifestyle_dat4)

Anova(logreg_lifestyle_sex1)
summary(logreg_lifestyle_sex1)
```

#Formatting output using Stargazer

```{r}

stargazer(logreg_lifestyle_sex1, type = "text", title = "Multinomial Regression Results", dep.var.labels = c("Underweight", "Overweight","Obese"),omit.stat = c("LL", "ser","f"), no.space = TRUE, align=TRUE, model.names = T)

```

#Export table
```{r}
#tab_model(logreg_lifestyle_sex1)
tab_model(logreg_lifestyle_sex1,   show.se = TRUE, show.p= TRUE, show.std = TRUE, show.stat = TRUE)
```


#There is a significant interaction between sex and other variables except race and sedentary behaviour, hence, we Stratify by sex using the "subset ()" in R to see the effect of the interaction across variables

```{r}
#male
fit_male<- multinom( BMIGroups1~FoodChoice + SleepHrs + SedentaryBehaviour+  Race + Decade,  data = lifestyle_dat4, subset = Sex=="Male")
```

```{r}
#female
fit_female<- multinom( BMIGroups1~FoodChoice + SleepHrs + SedentaryBehaviour+  Race + Decade,  data = lifestyle_dat4, subset = Sex=="Female")

```
```{r}
tab_model(fit_male,  p.style = "stars", title = "Male Weight Variation" )
```


```{r}
tab_model(fit_female, p.style = "stars", title = "Female Weight Variation" )
```

#Present all model coefficients next to one another to see the effect of interaction
```{r}
fit_finalmodel = stargazer(fit_male, fit_female, type = "text", title = "Regression Results", dep.var.labels = c("Underweight", "Overweight","Obese"),omit.stat = c("LL", "ser","f"), no.space = TRUE
          , align=TRUE, 
          model.names = T, column.labels = c("Male","Male", "Male", "Female", "Female", "Female"))
```


#Calculate Z score and p-value for the variables in the multinomial logistic regression model

```{r Calculate z statistics}
z_stats <- summary(logreg_lifestyle_sex1)$coefficients/summary(logreg_lifestyle_sex1)$standard.errors

stargazer(z_stats, type = "text", title = "Z score of variables")

#display z_stats in transposed data frame
data.frame(t(z_stats))
```

#Calculate the p-value for each variable

```{r}
p_value <- (1-pnorm(abs(z_stats), 0, 1))*2


stargazer(p_value, type = "text", title = "P-values of variables")

#display p-values in transposed data frame
data.frame(t(p_value))

```

#Check for fitted values

```{r}
#head(fitted(logreg_lifestyle_sex1))
fitvalues = fitted(logreg_lifestyle_sex1)

head(fitvalues)
```
#Calculate the Relative Risk

```{r}
rrisk <- exp(coef(logreg_lifestyle_sex1))

#display relative risk values in transposed data frame

data.frame(t(rrisk))

```


#Calculate Confidence Interval, Odds

```{r}
#calculate confidence interval
coef02 = coef(logreg_lifestyle_sex1)
CI = round(confint(logreg_lifestyle_sex1),2)
CI


```

```{r}
#Calculate odds ratio
OR_lifestyle = exp(cbind(OR=coef02, CI))
OR_lifestyle

#display odds ratio values in transposed data frame

data.frame(t(OR_lifestyle))


```
Here we see the odds ratio and relative risk have the same values




#Model Evaluation

```{r}
pred = predict(logreg_lifestyle,type='probs')

head (pred)
```
This returns the  predicted probabilities for each class.

```{r}
pred1 = predict(logreg_lifestyle,type='class')

pred1
```

This returns the  classication levels for each class.

```{r}
pred1 <- predict(logreg_lifestyle,type='class')
set.seed(1)
s <- sample(1:100)

pred1[s]
```

```{r}

# Using sample_frac to create 70 - 30 split into test and train
train <- sample_frac(lifestyle_dat4, 0.7)
sample_id <- as.numeric(rownames(train)) # rownames() returns character so as.numeric
test <- lifestyle_dat4[-sample_id,]

# Building classification table
ctable <- table(lifestyle_dat4$BMIGroups1, pred1)
 
# Calculating accuracy - sum of diagonal elements divided by total obs
round((sum(diag(ctable))/sum(ctable))*100,2)
```
```{r}
test$precticed <- predict(logreg_lifestyle, newdata = test, "class")
 
# Building classification table
ctable <- table(test$BMIGroups1, test$precticed)
 
# Calculating accuracy - sum of diagonal elements divided by total obs
round((sum(diag(ctable))/sum(ctable))*100,2)
```

The accuracy of the test dataset is almost same,  compared to training dataset.

The prediction accuracy of our model is 50%, 


```{r}
#Getting the ROC curve
pred = predict.glm(logreg_lifestyle,type='response')
roc01 = multiclass.roc(lifestyle_dat4$BMIGroups1,pred)
roc01


```
```{r}
VarInfFact = vif(logreg_lifestyle_sex)
VarInfFact1 = round(VarInfFact, 2)
```

